---
title: "Basic inferential data analysis"
author: "Georgy Makarov"
date: "April 24, 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

## Overview

This is the part two of the course project of Statistical Inference course. Here 
we use confidence intervals and hypothesis tests to compare tooth growth by 
supplement and dose from the *ToothGrowth* data. There are supporting plots in 
appendix to this project.

## Summary of the data

ToothGrowth dataset shows the length of cells responsible for tooth growth in 60 observations. There are three features in the dataset: *len* stands for the tooth 
length; *supp* is a factor with supplement *VC* or *OJ*; *dose* shows the dose
of the supplement received.

```{r}
data("ToothGrowth")
summary(ToothGrowth)
```

*Dose* feature has three levels: 0.5, 1.0, 2.0. Therefore we consider it a
factor variable.

```{r}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
unique(ToothGrowth$dose)
```

The distribution of the tooth length is slightly skewed to the left, as the
mean is lower than the median. The histogram of the tooth length is in appendix
in *Figure 1*.

While the dose is 2.0 mg there is no difference in the impact of delivery method
on the tooth growth. When the dose is either 0.5 or 1.0 mg the orange juice is 
more effective for the tooth growth. Visualization of this idea is given in 
appendix in *Figure 2*.

## Hypothesis testing

We test three hypothesis. First hypothesis: tooth growth does not depend on 
application method. Second hypothesis: tooth growth depends on application method
when the dose is small. Third hypothesis: tooth growth does not depend on the
dose of vitamin C. These hypothesis appear through $H_0$ notation:

1. H~0~: OJ and VC means are the same;  
2. H~0~: OJ and VC means are the same when dose is lower than 2;  
3. H~0~: 0.5 and 2.0 doses means are the same.

### Hypothesis 1

```{r}
t.test(len ~ supp, data = ToothGrowth)
```

P-value is **0.06** and it is greater than 0.05. Confidence interval includes 0. 
There is no evidence to reject $H_0$ - application methods do not impact tooth
growth.

### Hypothesis 2

```{r}
t.test(len ~ supp, data = ToothGrowth[ToothGrowth$dose %in% c(0.5, 1.0),])
```

P-value is **0.004**, which is less than 0.05. Confidence interval does not 
contain 0. There is evidence enough to reject $H_0$ - application methods impact 
tooth growth when the dose is less than 2.0 mg.

### Hypothesis 3

```{r}
t.test(len ~ dose, data = ToothGrowth[ToothGrowth$dose %in% c(0.5, 2),])
```

P-value is **4.4e-14**. which is less than 0.05. Confidence interval does not 
contain 0. There is evidence enough to reject $H_0$ - the dose impacts on tooth 
growth.

## Conclusions

This analysis assumed that each subject was randomly assigned and it was a 
representative of the population. It is also based on the assumption that all
observations are independent. The analysis shows that the tooth growth depends 
on the dose of vitamin C. While the method of application is critical when the 
dose is below 2.0 mg, it does not imply any difference on tooth growth when 
the dose icreases.

## Appendix

```{r, echo=FALSE,  out.width='75%', out.height='75%'}
hist(ToothGrowth$len, breaks = 5, xlim = c(0, 35), main = "Tooth growth",
     col = "skyblue", xlab = "Tooth length")
```

Figure 1. Distribution of tooth length

```{r, echo=FALSE,  out.width='75%', out.height='75%'}
boxplot(len ~ supp*dose, data = ToothGrowth, col = (c("gold", "skyblue")), 
        xlab = "Supplement and growth", main = "Tooth growth", ylab = "Length")
```

Figure 2. Tooth growth vs supplement & dose







